High resolution, annual maps of the characteristics of smallholder-dominated croplands at national scales

### 2021-06-28 inst/manuscript/figures.Rmd
Revised through image compositing methods, removed figure 2 to SI, reorganized figures.Rmd to accommodate.

# database results
load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))
```

@@

### Figure 2: Grid system and AOIs

```{r, eval = FALSE}
ghana_aois <- tiles %>% pull(aoi1) %>% unique
ghana_degrees <- tiles %>% pull(aoi2) %>% unique

ind <- 8  # index of selected AOI for plotting (from ghana_aois)
xylims <- st_coordinates(ghana)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))

@@

gp <- cowplot::plot_grid(p1, p2, p3, nrow = 1, scale = c(1, 1, 0.7), 
                         labels = LETTERS[1:3], label_x = 0.05)
ggsave(pth("figures/figure2.png"), width = 7, height = 3,
       units = "in", dpi = 300)
```

### Figure 3: Labeller interface
Taken as a screenshot.

### Figure S1: Bayesian Risk

@@

### Figure S2: Map reference sample
```{r, eval = FALSE}
ref_sample <- readr::read_csv(
  system.file("extdata", "map_reference_sample.csv", package = "activemapper")
)
ref_sample <- ref_sample %>% st_as_sf(coords = c("x", "y"), crs = 4326)

@@

## Results 
### Figure 4: Composite quality
```{r, eval=FALSE}
data("image_quality")

@@

### Figure S3: PlanetScope image density
The density of imagery available for download from the image catalog.

@@

### Figure S4: Initial training site distribution
Figure showing distribution of initial 500 samples per cluster, as well as distribution of points per iteration.  
```{r trainingsites, eval=FALSE}
data("train_val_sites")

@@

### Figure S5: Label count and quality
```{r labelquality, eval=FALSE}
data("label_summary")

@@

### Figures S6 and S7: Consensus label (Bayesian) risk

@@

### Figure S8: Random versus active sampling
```{r, eval=FALSE}
data("iteration_metrics")

@@

### Figure S9: Training data error impact on probability images
```{r, eval = FALSE}
probpal <- viridis::viridis
cprob <- raster(here::here("external/prob/image_c294_r532.tif"))
hprob <- raster(here::here("external/prob/image_c294_r532_high.tif"))
lprob <- raster(here::here("external/prob/image_c294_r532_low.tif"))

@@

### Figure S10: AOI Zones

@@

### Figures S11 and S12: Segmentation statistics (grid scale)

@@

### Figure S13: Average fields sizes

>>>>

# database results
load(system.file("extdata/", "instance_tbls.rda", package = "activemapper"))

# Ghana data
ghana_aois <- tiles %>% pull(aoi1) %>% unique
ghana_degrees <- tiles %>% pull(aoi2) %>% unique
```

@@

### Figure 2: Labeller interface
Taken as a screenshot.

### Figure S1: Grid reference system

```{r, eval = FALSE}
ind <- 8  # index of selected AOI for plotting (from ghana_aois)
xylims <- st_coordinates(ghana)
maprange <-  c(range(xylims[, 1]), range(xylims[, 2]))

@@

gp <- cowplot::plot_grid(p1, p2, p3, nrow = 1, scale = c(1, 1, 0.7), 
                         labels = LETTERS[1:3], label_x = 0.05)
ggsave(pth("figures/reference_system.png"), width = 7, height = 3,
       units = "in", dpi = 300)
```

### Figure S2: Bayesian Risk

@@

### Figure S3: Map reference sample
```{r, eval = FALSE}
ref_sample <- readr::read_csv(
  system.file("extdata", "map_reference_sample.csv", package = "activemapper")
)
ref_sample <- ref_sample %>% st_as_sf(coords = c("x", "y"), crs = 4326)

@@

## Results 
### Figure 3: Composite quality
```{r, eval=FALSE}
data("image_quality")

@@

### Figure S4: PlanetScope image density
The density of imagery available for download from the image catalog.

@@

### Figure S5: Initial training site distribution
Figure showing distribution of initial 500 samples per cluster, as well as distribution of points per iteration.  
```{r trainingsites, eval=FALSE}
data("train_val_sites")

@@

### Figure S6: Label count and quality
```{r labelquality, eval=FALSE}
data("label_summary")

@@

### Figures S7 and S8: Consensus label (Bayesian) risk

@@

### Figure S9: Random versus active sampling
```{r, eval=FALSE}
data("iteration_metrics")

@@

### Figure S10: Training data error impact on probability images
```{r, eval = FALSE}
probpal <- viridis::viridis
cprob <- raster(here::here("external/prob/image_c294_r532.tif"))
hprob <- raster(here::here("external/prob/image_c294_r532_high.tif"))
lprob <- raster(here::here("external/prob/image_c294_r532_low.tif"))

@@

### Figure S11: AOI Zones

@@

### Figures S12 and S13: Segmentation statistics (grid scale)

@@

### Figure S14: Average fields sizes

### 2021-06-28 inst/manuscript/manuscript.Rmd
Revised through image compositing methods, removed figure 2 to SI, reorganized figures.Rmd to accommodate.

Here we use this approach to create a high resolution, country-scale map of crop field boundaries in Ghana, a country where smallholder farming predominates, and which has a broad mix of climates and agricultural systems, including large areas where shifting agriculture is practiced [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016]. The map we develop represents a single agricultural year (2018-2019), as opposed to a multi-year composite, thereby demonstrating a capacity for annual, high resolution maps that can be used to monitor rapidly evolving small-scale agricultural systems, including key characteristics such as field size. In addition to presenting one of the most spatially extensive agricultural applications of CubeSats to date, our findings provide insights into how maps can be improved by adhering to best practices for model training [@ElmesAccountingtrainingdata2020].

@@

## System overview

The mapping system design has four primary components focused on 1) image acquisition and compositing, 2) training and validation data collection, 3) image classification, and 4) segmentation. The first component is applied as a once-off step to generate an image catalog covering the mapping geography. The second component provides tools for labelling imagery, and controls an interactive model training and classification (component 3) pipeline that produces a map of cropland probabilities for each image pixel (Figure \ref{fig:systemoverview}). The final segmentation step (component 4) is then initiated and applied to both the input image catalog and the posterior probability maps, resulting in vectorized field boundaries. Each system component comprises an individual software module designed to be implemented on cloud computing architecture, and available on a GitHub repository (see data and software availability section for details).

```{r systemoverview, echo = FALSE, out.width="50%", fig.cap="An overview of the primary system components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them. The dashed line indicates iterative interactions, while solid lines indicate one-time or irregular connections.", fig.align='center'}
knitr::include_graphics('figures/figure1.png')
```

## System components

### Image compositing

The image processing components of our system were designed to work with PlanetScope Analytic surface reflectance data [@planetteamPlanetApplicationProgram2018]. PlanetScope provides three visual (red, green, blue) and a near-infrared band at 3-4 m resolution at nominal daily frequency. Although these images are already pre-preprocessed and corrected for atmospheric effects, there are residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016]. To correct for these factors, we developed a procedure for creating temporal composites representing the primary growing and non-growing seasons within a single year.

PlanetScope imagery is accessed via the Planet API [@planetteamPlanetApplicationProgram2018], and an initial order is placed for all imagery falling within the mapping geography and the date ranges for the two compositing periods. The imagery is collected and transferred directly to a cloud storage platform (Amazon Web Services [AWS] S3).

Individual images are then transformed into analysis ready data (ARD) [@DwyerAnalysisReadyData2018], by subsetting each downloaded image into 0.05 degree tiles, regardless of cloud cover. Tiles are organized within a larger 1 degree resolution grid that covers the entire continent, which defines the minimum mapping area of interest (AOI; Figure \ref{fig:aois}).

<!-- and then applies the usable data mask (UDM) supplied with the imagery to remove low quality and cloudy pixels.  -->

```{r aois, echo = FALSE, out.width="100%", fig.cap="The reference system used in the mapping platform, including A) numbered areas of interest (AOIs) that define the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), B) the 0.05 degree tile used for compositing imagery, and C) the 0.005 degree resolution reference grid used for collecting training data and distributed computing.", fig.align='center'}
knitr::include_graphics('figures/figure2.png')
```

The temporal compositing process is applied to the tiled daily images for the time period of interest, which in this case was one of two multi-month seasons, the primary growing and dry seasons for a single agricultural year. Imagery from two seasons helps to improve the performance of cropland classifiers [@Debatsgeneralizedcomputervision2016], while having the seasons in the same year helps to minimize differences caused by land change. For each pixel in each image in each ARD temporal stack for a given season, two weights are calculated:

@@

```{=tex}
\begin{equation} \label{eq:shadow}  
\mathrm{W2_t} =\begin{cases}
    \frac{1}{\mathrm{NIR_t}^4}, & \text{if $\mathrm{NIR_t}$ < median\{$\mathrm{NIR_{t1}}$, $\mathrm{NIR_{t2}}$, ..., $\mathrm{NIR_{ti}}$\}}.\\
    1, & \text{otherwise}.
  \end{cases}
\end{equation}
```
Where *t* is a particular date in near-daily time series of PlanetScope images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band. Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to haze and cloud pixels [@Zhangimagetransformcharacterize2002], while Equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow considering the significant darkening effect of the cloud shadows in the Near Infrared band [@ZhuObjectbasedcloudcloud2012; @qiuCirrusCloudsThat2020]

Once these two weights are calculated, the final composited pixel value for each of the four PlanetScope bands is:

```{=tex}
\begin{equation}
\mathrm{\bar{B} = \frac{\sum_{t=1}^{T}B_t * W1_t * W2_t}{\sum_{t=1}^{T}W1_t * W2_t}}
\end{equation}
```
Which is a weighted mean for each pixel for each band *B* for the particular compositing period. The composited tiles were then added to the S3 store (Figure \ref{fig:systemoverview}), where they are stored as cloud-optimized geotiffs, and a "slippy map\footnote{https://wiki.openstreetmap.org/wiki/Slippy\_Map}" rendering is created for each composite using Raster Foundry [@azaveaRasterFoundry2020]. The web-rendered imagery is presented within the training data platform (next section).

### Labelling platform

>>>>

Here we use this approach to create a high resolution, country-scale map of crop field boundaries in Ghana, a country where smallholder farming predominates, and which has a broad mix of climates and agricultural systems, including large areas where shifting agriculture is practiced [@KansangaTraditionalagriculturetransition2019; @SambergSubnationaldistributionaverage2016]. The map we develop represents a single agricultural year (2018-2019), as opposed to a multi-year composite, thereby demonstrating a capacity for annual, high resolution maps that can be used to monitor rapidly evolving small-scale agricultural systems, including key characteristics such as field size. In addition to presenting one of the most spatially extensive agricultural applications of CubeSats to date, we provide a new technique for converting daily imagery into seasonal composites, and show how best practices for model training and label collection can be applied to improve map accuracy [@ElmesAccountingtrainingdata2020].

@@

The cropland mapping approach we developed is comprised of four primary components (Figure \ref{fig:systemoverview}) that were developed using open source software, and is designed to run in a cloud computing environment. The first component collects daily PlanetScope imagery collected over into cloud-free seasonal composites. The second component is a custom-built platform that a provides a set tools for labelling the PlanetScope composites, and built-in procedures for assessing and minimizing label error. This platform interacts with the third component, a machine learning process running in Apache Spark, which includes a routine for 1) selecting new training sites based on their prediction uncertainty, and 2) returning these to the labelling platform for labeling. This interaction, which represents active learning [@cohnImprovingGeneralizationActive1994; @tuiaSurveyActiveLearning2011], repeats until a model performance threshold is reached. The fourth and final component is a segmentation process that is applied to the seasonal image composites to segment the entire image, and then filter using the pixel-wise probability maps resulting from active learning, resulting in a final set of vectorized field boundaries. Each system component is available on an open repository (see data and software availability section for details).

```{r systemoverview, echo = FALSE, out.width="50%", fig.cap="An overview of the primary mapping components, the data stores that hold the inputs and outputs from each component, and the direction of connections between them. The dashed line indicates iterative interactions, while solid lines indicate one-time or irregular connections.", fig.align='center'}
knitr::include_graphics('figures/figure1.png')
```

## Image compositing

The image processing component was designed for PlanetScope Analytic surface reflectance imagery [@planetteamPlanetApplicationProgram2018], which provides three visual (red, green, blue) and near-infrared bands at 3.7 m resolution at nominal daily frequency. The images are provided as ortho-rectified and converted to surface reflectance, although there are residual errors from inter-sensor differences and the radiometric normalization process [@HouborgDailyRetrievalNDVI2018], variation in the orientation of scene footprints, as well as a high frequency of cloud cover over the study region [@WilsonRemotelySensedHighResolution2016] that are not fully captured by the provided cloud masks. To minimize the effect of these residual errors, we developed a procedure for creating temporal composites of the primary growing and non-growing seasons within a single 12-month period. For Ghana, we defined the primary growing season as May through September, followed by the off (or dry) season from November or December through February.
We chose these two seasons because prior work shows that the contrast improves cropland classifications [@Debatsgeneralizedcomputervision2016], while having the seasons captured in sequence in the same year helps minimize differences caused by land change. The wide time intervals we used to define each season were necessary for collecting a sufficient number of images to make high quality composites, as Ghana's cloud cover renders many scenes unusable (and thus unavailable in Planet's catalog), thus the frequency of available imagery can be substantially less than daily during several months.

We collected all available scenes intersecting Ghana and falling within these two seasons during the 2018 agricultural year (beginning in May and ending in February, 2019) via the Planet API [@planetteamPlanetApplicationProgram2018], and transferred these to cloud storage (Amazon Web Services [AWS] S3). We then converted each scene into analysis ready  data [@DwyerAnalysisReadyData2018], by cropping each downloaded to the dimensions of a 0.05 degree grid (see Figure S1 in Supplemental Information [SI]), which provided the dimensions for our composited image tiles. We chose this dimension for tiling because it  is slightly narrower than the short axis of a PlanetScope scene, which increases the number of intersecting scenes that completely cover the tile, helping to minimize edge artifacts in the composites.

To create a composite tile for a given season, we calculated two weights for each pixel in each image within the ARD stack for that season:

@@

```{=tex}
\begin{equation} \label{eq:shadow}  
\mathrm{W2_t} =\begin{cases}
    \frac{1}{\mathrm{NIR_t}^4}, & \text{if $\mathrm{NIR_t}$ < median\{$\mathrm{NIR_{t1}}$, $\mathrm{NIR_{t2}}$, ..., $\mathrm{NIR_{ti}}$\}}.\\
    1, & \text{otherwise}.
  \end{cases}
\end{equation}
```
Where *t* is a particular date in the time series of images, which begins at date 1 for the given compositing period and ends on date *i*, *blue* is the blue band, and *NIR* the near infrared band. Equation \ref{eq:cloud} assigns lower weights to hazy and clouded pixels as the blue band is sensitive to haze and cloud pixels [@Zhangimagetransformcharacterize2002], while Equation \ref{eq:shadow} assigns low weights to pixels in cloud shadow considering the significant darkening effect of the cloud shadows in the Near Infrared band [@ZhuObjectbasedcloudcloud2012; @qiuCirrusCloudsThat2020]

After assigning these two weights, we calculated the final composited pixel value:

```{=tex}
\begin{equation}
\mathrm{\bar{B} = \frac{\sum_{t=1}^{T}B_t * W1_t * W2_t}{\sum_{t=1}^{T}W1_t * W2_t}}
\end{equation}
```
Which is the weighted mean for each pixel for each band *B* for the particular compositing period. The composited tiles were then saved as cloud-optimized geotiffs, and a "slippy map\footnote{https://wiki.openstreetmap.org/wiki/Slippy\_Map}" rendering was created for each composite using Raster Foundry [@azaveaRasterFoundry2020], for display within the labelling platform (next section).

## Labelling platform

### 2021-06-28 inst/manuscript/si.Rmd
Revised through image compositing methods, removed figure 2 to SI, reorganized figures.Rmd to accommodate.

# Supplemental Methods
## Assessing quality of composites
The rubric presented in Table \ref{tab:imqcriteria} was used to assess the quality of the seasonal image composites. The imagery was evaluated by examining their Raster Foundry [@azaveaRasterFoundry2020] overlays within a `labeller` instance set up for the purpose.

>>>>

# Supplemental Methods
## Map reference system
The map reference system used in our mapping approach had three different levels. At the coarsest level, Ghana was divided into 16 different mapping zones, or areas of interest (AOIS; Figure \ref{fig:aois}A), that were used to create AOI-specific mapping models. Each AOIs was comprised of 400 to 777 adjacent tiles. These tiles were used to create seasonal image composites, and were defined within a 0.05 degree grid (Figure \ref{fig:aois}B)), with each tile numbered to correspond to a larger 1 degree grid cell that it sits within (dotted lines in Figure \ref{fig:aois}B). Training and reference labels were created within a 0.005 degree grid that nested within each tile (Figure \ref{fig:aois}C). Therefore, each tile has 100 grid cells, and there are 400 tiles per 1X1 degree. The smallest AOI consists of a single 1X1 degree, which fall in the center of the country (AOIs 5, 8, 11, and 14). AOIs falling along Ghana's boundaries were created by tiles from 1X1 degree cells that straddled Ghana's border with those from the closest degree cells that were fully contained within Ghana (e.g. AOI 1), with the exception of AOI 16, which was comprised of tiles in three partial 1X1 degree cells along Ghana's coast.

```{r aois, echo = FALSE, out.width="100%", fig.cap="The reference system used in this mapping approach, including A) numbered areas of interest (AOIs) that define the minimum mapping geography (solid black lines; dotted lines indicate boundaries of 1 degree grid), B) the 0.05 degree tile used for compositing imagery, and C) the 0.005 degree resolution reference grid used for collecting training data and distributed computing.", fig.align='center'}
knitr::include_graphics('figures/reference_system.png')
```

## Composites Assessing quality of composites
The rubric presented in Table \ref{tab:imqcriteria} was used to assess the quality of the seasonal image composites. The imagery was evaluated by examining their Raster Foundry [@azaveaRasterFoundry2020] overlays within a `labeller` instance set up for the purpose.
